---
title: "Lecture 2"
author: "Lakshya Agarwal"
date: "9/02/2023"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: true
editor: 
  markdown: 
    wrap: 72
---

# Lab 1

```{r}
require(visreg)
require(glue)
```

## Importing data

```{r}
data <- read.csv('./video_games_fall_2023.csv')
attach(data)
head(data)
```

## 1. Visualizing variables

### `score`

```{r}
summary(score)

boxplot(score, col="blue", ylab="Score", main="Boxplot of score")

hist(score, breaks=20, col="blue")
```

### `sales_global`

```{r}
summary(sales_global)

boxplot(sales_global, col="green", ylab="Global sales (in millions)", main="Boxplot of sales_global")

hist(sales_global, breaks=100, col="green")
```

### `release_year`

```{r}
summary(release_year)

boxplot(release_year, col="orange", ylab="Release Year", main="Boxplot of release_year")

hist(release_year, breaks=25, col="orange")

```

### `count_critic`

```{r}
summary(count_critic)

boxplot(count_critic, col="purple", ylab="# of critics", main="Boxplot of count_critic")

hist(count_critic, breaks=10, col="purple")
```

### Scatterplots

```{r}
par(mfrow = c(1,3))
plot(sales_global, score, col=factor(genre), main = "sales_global", pch=19)
legend("bottomright", legend = levels(factor(genre)), col=factor(levels(factor(genre))), pch=19)
plot(release_year, score, col=factor(genre), main = "release_year", pch=19)
plot(count_critic, score, col=factor(genre), main = "count_critic", pch=19)
```

## 2. Simple linear regression

### `score = bo + b1(sales_global)`

```{r}
model_1 = lm(score ~ sales_global)
summary(model_1)

confint(model_1, "sales_global", level = 0.95)

visreg(model_1, alpha = 0.05, main=glue("score ~ sales_global (R-squared: {format(summary(model_1)$r.squared, digits = 2)})"))
```

### `score = bo + b1(release_year)`

```{r}
model_2 = lm(score ~ release_year)
summary(model_2)

confint(model_2, "release_year", level = 0.95)

visreg(model_2, alpha = 0.05, main=glue("score ~ release_year (R-squared: {format(summary(model_2)$r.squared, digits = 2)})"))
```

### `score = bo + b1(count_critic)`

```{r}
model_3 = lm(score ~ count_critic)
summary(model_3)

confint(model_3, "count_critic", level = 0.95)

visreg(model_3, alpha = 0.05, main=glue("score ~ count_critic (R-squared: {format(summary(model_3)$r.squared, digits = 2)})"))
```

## 3. Predictions

```{r}
obs_1 <- data.frame(sales_global = 0.75)
predict(model_1, newdata = obs_1)

obs_2 <- data.frame(release_year = 2009)
predict(model_2, newdata = obs_2)

obs_3 <- data.frame(count_critic = 80)
predict(model_3, newdata = obs_3)
```

## 4. Multiple regression
### `score = bo+ b1(sales_global) + b2(release_year) + b3(count_critic)`

```{r}
model_4 <- lm(score ~ sales_global + release_year + count_critic)
summary(model_4)

obs_4 <- data.frame(sales_global = 0.75, release_year = 2009, count_critic = 80)
predict(model_4, newdata = obs_4)
```

## 5. Categorical variables

### `score = bo+ b1(release_year) + b2(Nintendo)`

```{r}
data$nintendo <- as.factor(ifelse(data$publisher == "Nintendo", 1, 0))
attach(data)
levels(data$nintendo)
table(data$nintendo)
```
```{r}
model_5 <- lm(score ~ release_year + nintendo)
summary(model_5)
```

```{r}
b0 = coef(model_5)[1]
b1 = coef(model_5)[2]
b2 = coef(model_5)[3]

plot(release_year, score, pch=0, col=ifelse(nintendo == 1, "green", "blue"), main = glue("score ~ release_year + nintendo (R-squared: {format(summary(model_5)$r.squared, digits = 2)})"))
abline(b0+b2, b1, col="green", lty="dashed", lwd=2)
abline(b0, b1, col="blue",  lty="dashed", lwd=2)
legend("bottomleft", pch=0, col = c('green', 'blue'), legend = c('Nintendo', 'Non-Nintendo'))
```

## 6. Categorical variable with multiple categories

### `score = bo+ b1(genre)`
```{r}
table(genre)
```

```{r}
data$genre = relevel(as.factor(data$genre), ref="Racing")
attach(data)

model_6 <- lm(score ~ genre)
summary(model_6)
```
```{r}
data$genre = relevel(as.factor(data$genre), ref="Shooter")
attach(data)

model_7 <- lm(score ~ genre)
summary(model_7)

```

## 7. Interaction terms
### `score = bo+ b1(Nintendo)+ b2(strategy)+b3(strategy*Nintendo)`
```{r}
data$strategy = as.factor(ifelse(data$genre == "Strategy", 1, 0))
attach(data)
```

```{r}
model_8 <- lm(score ~ nintendo + strategy + nintendo*strategy)
summary(model_8)
```

### `score = bo+ b1(release_year) + b2(Nintendo) + b3(release_year*Nintendo)`

```{r}
model_9 = lm(score ~ release_year + nintendo + release_year*nintendo)
summary(model_9)

b0 = coef(model_9)[1]
b1 = coef(model_9)[2]
b2 = coef(model_9)[3]
b3 = coef(model_9)[4]

plot(release_year, score, pch=2, col=ifelse(nintendo == 1, "green", "blue"), main = glue("score ~ release_year + nintendo + 
                                                                                         release_year*nintendo (R-squared: {format(summary(model_5)$r.squared, digits = 2)})"))
abline(b0+b2, b1+b3, col="green", lty="dashed", lwd=2)
abline(b0, b1, col="blue",  lty="dashed", lwd=2)
legend("bottomleft", pch=2, col = c('green', 'blue'), legend = c('Nintendo', 'Non-Nintendo'))

```

