---
title: "MGSC 661-075 - Lab 2"
output: html_notebook
author: "Lakshya Agarwal, Jatin Suri"
---

# Lab 2

```{r}
require(visreg)
require(glue)
require(car)
require(lmtest)
require(splines)
require(psych)
require(stargazer)
require(plm)
require(ggplot2)
require(ggpubr)
require(caTools)
require(methods)
require(boot)
```

## Importing data

```{r}
data <- read.csv("./board_games_fall_2023.csv")
attach(data)
head(data)
```

## 1. Model Issues: Non-linearity

#### `avg_rating = bo+ b1(year)+ b2(avg_timeplay)+ b3(weight)`

#### Part A:

```{r}
reg1 <- lm(avg_rating ~ year + avg_timeplay + weight)
residualPlots(reg1)
```

## 2. Model Issues: Heteroskedasticity

#### `avg_rating = bo+ b1(avg_timeplay)`

#### Part B:

```{r}
reg2 <- lm(avg_rating ~ avg_timeplay)
summary(reg2)
```

#### Part C:

```{r}
residualPlot(reg2, quadratic = FALSE)
```

#### Part D:

```{r}
ncvTest(reg2)
```

#### Part E:

```{r}
coeftest(reg2, vcov = vcovHC(reg2, type = "HC1"))
```

## 3. Model Issues: Outliers

#### `avg_rating = bo+ b1(min_players)+ b2(age)+ b3(num_votes)`

#### Part A:

```{r}
reg3 <- lm(avg_rating ~ min_players + age + num_votes)
summary(reg3)
qqPlot(reg3, envelope = FALSE)
outlierTest(reg3)
```

#### Part B:

```{r}
data[3124, ]
```

#### Part C:

```{r}
new_data <- data[-c(3124), ]

reg4 <- lm(data = new_data, formula = avg_rating ~ min_players + age + num_votes)
summary(reg4)
```

## 4. Model Issues: Collinearity

#### `avg_rating = bo+ b1(year)+ b2(age)+ b3(min_timeplay)+ b4(max_timeplay)`

#### Part B:

```{r}
quantvars <- data[, c(7, 8, 9, 12)]
corr_matrix <- cor(quantvars)
round(corr_matrix, 2)
```

#### Part D:

```{r}
reg5 <- lm(avg_rating ~ year + age + min_timeplay + max_timeplay)
summary(reg5)
vif(reg5)
```

#### Part F:
```{r}
reg5_1 <- lm(avg_rating ~ year + age + min_timeplay)
summary(reg5_1)
vif(reg5_1)
```


## 5. Presenting professional regression tables

#### Models
`reg1: avg_rating = bo+ b1(avg_timeplay)`

`reg2: avg_rating = bo+ b2(min_players)`

`reg3: avg_rating = bo+ b3(max_players)`

`mreg: avg_rating = bo+ b1(avg_timeplay)+ b2(min_players)+ b3(max_players)`

#### Part A:

```{r}
reg1 <- lm(avg_rating ~ avg_timeplay)
reg2 <- lm(avg_rating ~ min_players)
reg3 <- lm(avg_rating ~ max_players)
mreg <- lm(avg_rating ~ avg_timeplay + min_players + max_players)
stargazer(reg1, reg2, reg3, mreg, type = "html")
```

#### Part B:

```{r}
summary(reg2)
```

#### Part C:

```{r}
stargazer(reg1, reg2, reg3, mreg, 
          title = "Regression Results", 
          type = "html", 
          align = TRUE, 
          dep.var.labels = c("Board Game Rating"), 
          covariate.labels = c("Average timeplay", "Minimum number of players", "Maximum number of players"))
```

#### Part D:

```{r}
stargazer(reg1, reg2, reg3, mreg, 
          title = "Regression Results", 
          type = "html", 
          align = TRUE, 
          dep.var.labels = c("Board Game Rating"), 
          covariate.labels = c("Average timeplay", "Minimum number of players", "Maximum number of players"),
          digits = 2)
```
## 6. Polynomial regression: Age

`avg_rating = f(age)`

#### Part A:
```{r}
poly_reg_1 <- lm(avg_rating ~ age)
poly_reg_2 <- lm(avg_rating ~ poly(age, degree = 2))
poly_reg_3 <- lm(avg_rating ~ poly(age, degree = 3))
poly_reg_4 <- lm(avg_rating ~ poly(age, degree = 4))

names(poly_reg_1$coefficients) <- c("Constant", "Age")
names(poly_reg_2$coefficients) <- c("Constant", "Age", "Age^2")
names(poly_reg_3$coefficients) <- c("Constant", "Age", "Age^2", "Age^3")
names(poly_reg_4$coefficients) <- c("Constant", "Age", "Age^2", "Age^3", "Age^4")

stargazer(poly_reg_1, poly_reg_2, poly_reg_3, poly_reg_4,
          title = "Regression Results", 
          type = "html", 
          align = TRUE, 
          dep.var.labels = c("Board Game Rating"), 
          covariate.labels = c("Constant", "Age", "Age<sup>2</sup>", "Age<sup>3</sup>", "Age<sup>4</sup>"),
          column.labels = c("Linear (d=1)", "Quadratic (d=2)", "Cubic (d=3)", "Quartic (d=4)"),
          digits = 2)
```
#### Part B:
```{r fig.width=9, fig.height=7}
plot_dn <- function(data, n) {
  plot <- ggplot(data, aes(x = age, y=avg_rating)) +
    geom_point(color="grey") +
    geom_smooth(method = "lm", formula = if(n == 1) y ~ x else (y ~ poly(x, n)), aes(colour="blue"), se=FALSE) +
    labs(x = "Age", y = "Average rating", title = glue("Polynomial regression (d={n})"), caption = "Source: Board Game Data") +
    scale_color_identity(name = "Degree", breaks = c("blue"), labels=glue("d-{n}"), guide="legend") +
    theme_classic()
          
  return(plot)
}

plot_d1 <- plot_dn(data, 1)
plot_d2 <- plot_dn(data, 2)
plot_d3 <- plot_dn(data, 3)
plot_d4 <- plot_dn(data, 4)

ggarrange(plot_d1, plot_d2, plot_d3, plot_d4, nrow = 2, ncol = 2, legend = "top")

```

#### Part C:
```{r}
anova(poly_reg_1, poly_reg_2, poly_reg_3, poly_reg_4)
```

## 7. Polynomial regression: avg_timeplay

`avg_rating = f(avg_timeplay)`

#### Part A:
```{r}
poly_reg_1 <- lm(avg_rating ~ avg_timeplay)
poly_reg_2 <- lm(avg_rating ~ poly(avg_timeplay, degree = 2))
poly_reg_3 <- lm(avg_rating ~ poly(avg_timeplay, degree = 3))
poly_reg_4 <- lm(avg_rating ~ poly(avg_timeplay, degree = 4))

names(poly_reg_1$coefficients) <- c("Constant", "X")
names(poly_reg_2$coefficients) <- c("Constant", "X", "X^2")
names(poly_reg_3$coefficients) <- c("Constant", "X", "X^2", "X^3")
names(poly_reg_4$coefficients) <- c("Constant", "X", "X^2", "X^3", "X^4")


stargazer(poly_reg_1, poly_reg_2, poly_reg_3, poly_reg_4,
          title = "Regression Results (avg\\_rating ~ avg\\_timeplay)", 
          type = "html", 
          align = TRUE, 
          dep.var.labels = c("Board Game Rating"), 
          covariate.labels = c("Constant", "x", "x<sup>2</sup>", "x<sup>3</sup>", "x<sup>4</sup>"),
          column.labels = c("Linear (d=1)", "Quadratic (d=2)", "Cubic (d=3)", "Quartic (d=4)"),
          digits = 2)
```

#### Part B:
```{r, fig.width=8, fig.height=8}
plot_dn <- function(data, n) {
  plot <- ggplot(data, aes(x = avg_timeplay, y=avg_rating)) +
    geom_point(color="grey") +
    geom_smooth(method = "lm", formula = if(n == 1) y ~ x else (y ~ poly(x, n)), aes(colour="green"), se=FALSE) +
    labs(x = "Average Timeplay", y = "Average rating", title = glue("Polynomial regression (d={n})"), caption = "Source: Board Game Data") +
    scale_color_identity(name = "Degree", breaks = c("green"), labels=glue("d-{n}"), guide="legend") +
    theme_classic()
          
  return(plot)
}

plot_d1 <- plot_dn(data, 1)
plot_d2 <- plot_dn(data, 2)
plot_d3 <- plot_dn(data, 3)
plot_d4 <- plot_dn(data, 4)

ggarrange(plot_d1, plot_d2, plot_d3, plot_d4, nrow = 4, ncol = 1, legend = "right")
```

#### Part C:
```{r}
anova(poly_reg_1, poly_reg_2, poly_reg_3, poly_reg_4)
```

## 8. Multiple polynomial regression 

`avg_rating = f(age, avg_timeplay)`

#### Part A:
```{r}
poly_reg_final <- lm(avg_rating ~ avg_timeplay + poly(age, 3))
summary(poly_reg_final)
```

#### Part B:
```{r}
stargazer(poly_reg_final,
          title = "Regression Results", 
          type = "html", 
          align = TRUE, 
          dep.var.labels = c("Board Game Rating"), 
          covariate.labels = c("Average Timeplay", "Age", "Age<sup>2</sup>", "Age<sup>3</sup>", "Constant"),
          digits = 2)

```

## 9. Spline regression

`avg_rating = f(avg_timeplay)`

#### Part B:
```{r}
knots = quantile(avg_timeplay, probs = c(0.25, 0.50, 0.75))
spline_reg_1 <- lm(avg_rating ~ bs(avg_timeplay, knots = knots, degree = 1))
spline_reg_2 <- lm(avg_rating ~ bs(avg_timeplay, knots = knots, degree = 2))
spline_reg_3 <- lm(avg_rating ~ bs(avg_timeplay, knots = knots, degree = 3))


stargazer(spline_reg_1, spline_reg_2, spline_reg_3,
          title = "Spline Regression Results (avg\\_rating ~ avg\\_timeplay)", 
          type = "html", 
          align=TRUE,
          dep.var.labels = c("Board Game Rating"),
          column.labels = c("d=1", "d=2", "d=3"),
          digits = 2)
```


#### Part C:
```{r, fig.width=10, fig.height=7}
plot_spline_n <- function(data, n) {
  knots = quantile(avg_timeplay, probs = c(0.25, 0.50, 0.75))
  
  plot <- ggplot(data, aes(x = avg_timeplay, y = avg_rating)) +
  geom_point(color = "grey") +
  geom_smooth(
    method = "lm",
    formula = y ~ bs(x, knots = knots, degree = n),
    se = FALSE,
    aes(color = "red"),
  ) + 
  geom_vline(xintercept = knots, linetype="dotted", linewidth=0.75) +
  labs(x = "Average Timeplay", y = "Average Rating", title = glue("Spline regression (d={n})"), caption = "Source: Board Game Data") +
    scale_color_identity(name = "Spline", breaks = c("red"), labels=glue("Degree {n} - 3 knots"), guide="legend") +
    theme_classic()
  
  return(plot)
}

ggarrange(plot_spline_n(data, 1), plot_spline_n(data, 2), plot_spline_n(data, 3), ncol = 3, legend = "top")
```

## 10. Validation set test

`avg_rating = f(weight)`

### Part A:

```{r}
sample <- sample.split(data$avg_rating, SplitRatio = 0.5)

train_set <- subset(data, sample == TRUE)
test_set <- subset(data, sample == FALSE)

validation_mse <- function(degree, num_iter) {
  fit_dn <-
    lm(formula = if (degree == 1)
      avg_rating ~ weight
      else
        (avg_rating ~ poly(weight, degree))
      ,
      data = train_set)
  
  actual = test_set$avg_rating
  
  mses = rep(NA, num_iter)
  
  for (i in 1:num_iter) {
    prediction = predict(fit_dn, test_set)
    squared_error = (actual - prediction) ^ 2
    mses[i] = mean(squared_error)
  }
  
  mean(mses)
}

mse_degree = rep(NA, 10)

for (i in 1:10) {
  mse_degree[i] = validation_mse(i, 30)
  print(glue("d={i}: {format(mse_degree[i], digits=3)}"))
}

which.min(mse_degree)
min(mse_degree)
```

## 11. LOOCV test

`avg_rating = f(weight)`

### Part B:
```{r}
nrow(data)
```

### Part C:
```{r}
loocv_fit_d1 <- glm(formula = avg_rating ~ weight, data = data)
mse_loocv_fit_d1 <- cv.glm(data=data, loocv_fit_d1)$delta[1]

mse_loocv_fit_d1
```

### Part D:
```{r}
loocv_mse <- function(degree) {
  loocv_fit_dn <- glm(formula = if (degree == 1)
    avg_rating ~ weight
    else
      (avg_rating ~ poly(weight, degree))
    ,
    data = data)
  mse_loocv_fit_dn <- cv.glm(data = data, loocv_fit_dn)$delta[1]
  
  return (mse_loocv_fit_dn)
}

mse_loocv_list = rep(NA, 10)

for (i in 1:10) {
  mse_loocv_list[i] = loocv_mse(i)
  print(glue("d={i}: {format(mse_loocv_list[i], digits = 3)}"))
}

```

## 12. K-fold cross-validation

`avg_rating = f(num_votes)`

### Part B:
```{r}
k_fold_mse <- function(degree, k) {
  k_fold_fit_dn <- glm(formula = if (degree == 1)
    avg_rating ~ num_votes
    else
      (avg_rating ~ poly(num_votes, degree))
    ,
    data = data)
  
  mse_k_fold_fit_dn <- cv.glm(data = data, k_fold_fit_dn, K = k)$delta[1]
  
  return (mse_k_fold_fit_dn)
}

mse_k_fold_list <- rep(NA, 10)

for (i in 1:10) {
  mse_k_fold_list[i] = k_fold_mse(i, 5)
  print(glue("d={i}: {format(mse_k_fold_list[i], digits = 3)}"))
}

which.min(mse_k_fold_list)
min(mse_k_fold_list)
```

## 13. Cross-validation in a multiple spline model

`avg_rating = f(age,  year,  num_votes, avg_timeplay)`

### Part A:
```{r}
knots_age <- quantile(age, probs = c(0.25, 0.50, 0.75))
knots_year <- quantile(year, probs = c(0.25, 0.50, 0.75))
knots_num_votes <- quantile(num_votes, probs = c(0.25, 0.50, 0.75))
knots_avg_timeplay <- quantile(avg_timeplay, probs = c(0.25, 0.50, 0.75))

k_fold_spline <- function(a, b, c, d, k) {
  fit_model <- glm(
    formula =
      avg_rating ~
      bs(age, knots = knots_age, degree = a) +
      bs(year, knots = knots_year, degree = b) +
      bs(num_votes, knots = knots_num_votes, degree = c) +
      bs(avg_timeplay, knots = knots_avg_timeplay, degree = d)
  )
  
  mse <- cv.glm(data=data, fit_model, K=k)$delta[1]
  
  return (mse)
}
```

```{r}
start = Sys.time()

mse_spline_df = data.frame(
  a = rep(NA, 5 ^ 4),
  b = rep(NA, 5 ^ 4),
  c = rep(NA, 5 ^ 4),
  d = rep(NA, 5 ^ 4),
  mse = rep(NA, 5 ^ 4)
)


for (a in 1:5) {
  for (b in 1:5) {
    for (c in 1:5) {
      for (d in 1:5) {
        print(glue("Running 20-fold CV for a,b,c,d = {a},{b},{c},{d}"))
        
        mse_spline_df[nrow(mse_spline_df) + 1, ] <-
          c(a, b, c, d, k_fold_spline(a, b, c, d, 20))
      }
    }
  }
}
end = Sys.time()
```

```{r}
mse_spline_df[which.min(mse_spline_df$mse),]
```

