---
title: "MGSC 661-075 - Lab 4"
output: html_notebook
author: "Lakshya Agarwal, Jatin Suri"
---

# Lab 4

```{r}
require(car)
require(ggplot2)
require(bbplot)
require(ggtext)
require(methods)
require(caTools)
require(splines)
require(methods)
require(boot)
require(stargazer)
require(MASS)
require(klaR)
require(scales)
require(glue)
require(rms)
require(tree)
require(rpart)
require(rpart.plot)
require(randomForest)
require(gbm)
require(GGally)
require(ggfortify)

windowsFonts(Helvetica = "Product Sans")
```

```{r}
theme_lox <- function() {
  theme(
    panel.grid.major.x = element_line(size = 0.3, colour = "#cbcbcb"),
    panel.grid.major.y = element_line(size = 0.3, colour = "#cbcbcb"),
    plot.title = element_markdown(
      family = "Helvetica",
      size = 22,
      face = "bold",
      color = "#222222"
    ),
    plot.subtitle = element_text(
      family = "Helvetica",
      size = 16,
      margin = margin(2, 0, 2, 0)
    ),
    plot.caption = element_text(family = "Helvetica", face = "bold"),
    axis.text = element_text(
      family = "Helvetica",
      size = 12,
      color = "#222222"
    ),
    axis.title = element_text(
      family = "Helvetica",
      size = 14,
      color = "#222222"
    ),
    legend.text = element_text(family = "Helvetica", size = 12),
    legend.title = element_text(
      family = "Helvetica",
      size = 14,
      face = "bold"
    ),
    legend.position = "right",
    strip.text = element_text(
      family = "Helvetica",
      size = 12,
      hjust = 0.5
    )
  )
}
```

## Importing data

```{r}
diabetes_data <- read.csv("./diabetes - Copy.csv")
animal_shelter_data <- read.csv("./animal_shelter - Copy.csv")
animal_shelter_data$outcome = as.factor(animal_shelter_data$outcome)
animal_shelter_data$animal_type = as.factor(animal_shelter_data$animal_type)
animal_shelter_data$is_sterilized = as.factor(animal_shelter_data$is_sterilized)
animal_shelter_data$sex = as.factor(animal_shelter_data$sex)
attach(diabetes_data)
attach(animal_shelter_data)
```

## 1. Interpreting a classification tree

### Part C
```{r}
obs <- data.frame(x1 = c(8,4,4,2), x2=c(4,3,2,2), y=c(20,2,2,3))
obs
```

## 2. Growing a tree (outcome=f(age, number_animals_in_shelter, animal_type, is_sterilized, sex))

### Part A
```{r}
plot_tree <- function(cp) {
  tree <-
    rpart(
      outcome ~ age + number_animals_in_shelter + animal_type + is_sterilized + sex,
      data = animal_shelter_data,
      control = rpart.control(cp = cp)
    )
  
  rpart.plot(tree)
  
  return(tree)
}

tree_1 <- plot_tree(0.5)
tree_2 <- plot_tree(0.05)
tree_3 <- plot_tree(0.0005)
```

### Part C
```{r}
plotcp(tree_3)
```

### Part D
```{r}
optimal_cp <- tree_3$cptable[which.min(tree_3$cptable[, "xerror"]), "CP"]

optimal_cp
```
### Part E
```{r}
optimal_tree <- plot_tree(optimal_cp)

```

## 3. Growing a random forest

### Part F (outcome=f(age, number_animals_in_shelter, animal_type, is_sterilized, sex))
```{r}
rf <- randomForest(
  outcome ~ age + number_animals_in_shelter + animal_type + is_sterilized + sex,
  data = animal_shelter_data,
  ntree=1000, 
  importance=TRUE,
  na.action = na.omit
)

rf
```

### Part G
```{r}
rf_oob <- randomForest(
  outcome ~ age + number_animals_in_shelter + animal_type + is_sterilized + sex,
  data = animal_shelter_data,
  ntree=1000, 
  importance=TRUE,
  do.trace=100,
  na.action = na.omit
)

rf_oob
```
### Part I
```{r}
varImpPlot(rf_oob)
```

## 4. Boosting

### Part E (outcome=f(age, number_animals_in_shelter, animal_type, is_sterilized, sex))

```{r}
set.seed(420)
animal_shelter_data$outcome_category = ifelse(animal_shelter_data$outcome == "Adopted", 1, 0)
boosted_tree <- gbm(
  outcome_category ~ age + number_animals_in_shelter + animal_type + is_sterilized + sex,
  data=animal_shelter_data,
  distribution = 'bernoulli',
  n.trees = 1000,
  interaction.depth = 3
)

summary(boosted_tree)
```
```{r}
adoption_probs <- predict(boosted_tree, newdata = animal_shelter_data, type="response")
animal_shelter_data$predicted = ifelse(adoption_probs >= 0.5, 1, 0)
```

```{r}
classification_error <-
  1 - (
    sum(
      animal_shelter_data$predicted == animal_shelter_data$outcome_category
    ) / nrow(animal_shelter_data)
  )

classification_error
```

## 5. Principal Components Analysis

### Part C
```{r}
pca <- prcomp(diabetes_data[, c(1:8)], scale = TRUE)
pca
```

### Part D
```{r}
sort(abs(pca[2]$rotation[, 1]), decreasing = TRUE)
```

### Part E
```{r}
autoplot(
  pca,
  data=diabetes_data,
  loadings=TRUE,
  loadings.label=TRUE,
  color=ifelse(diabetes_data$outcome == 0, "red", "green")
)
```

